Linear Ordered Attribute Grammar scheduling using SAT-solving

### 2014-10-16 paper/paper.tex
Conclusion and discussion

\section{Conclusion and discussion} \label{sect:conclusion}
\Jeroen{todo: het werkt, misschien is andere representatie beter, en het is onvoorspelbaarder dus misschien duren sommige ags ineens heel lang om te schedulen}

>>>>

\section{Conclusion and discussion} \label{sect:conclusion}
We have explained the difficulties in attribute grammar scheduling and shown how to solve this problem using a SAT-solver. The given approach has been implemented and tested on a full-scale compiler built using attribute grammars. Results show that the algorithm works much faster than other approaches.

In the translation to the SAT-problem we have used a novel technique for ruling out cycles in the SAT problem using chordal graphs. This technique makes the problems much smaller than directly ruling out cycles, while encoding the same restrictions. We believe that this technique is applicable in other problems using SAT-solvers too.

Furthermore, we have shown that expressing the problem as a SAT problem has the advantage that extra constraints can be added. We illustrated this with two possible properties of the resulting schedule that an attribute grammar programmer may want to influence. Even though this makes the scheduling problem potentially harder, as the algorithm is left fewer choices, the solution is found very fast in all practical cases we have tried.

Another benefit of this approach that the attribute grammar scheduling can benefit from breakthroughs in the SAT community. The more efficient SAT solvers become, the better the attribute grammar scheduling becomes leading to larger and larger attribute grammars that are feasible to schedule.

One problem with the current approach in constrast to most other scheduling algorithms is the unpredictability. SAT-solvers use certain heuristics to quickly find solutions in many applications, but it can theoretically happen that for a certain attribute grammar the SAT problem that is generated is not suitable for these heuristics. A seemingly innocent change in the attribute grammar definition could therefore theoretically lead to a large increase in compile time. However, we have not encountered this problem and we believe that this situation is very unlikely to happen because of the maturity of the SAT-solvers.
